---
title: "Sobol-Assignment"
author: "Naomi Moraes"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

### Load libraries

```{r}
library(sensitivity)
library(tidyverse)
library(gridExtra)
library(purrr)
library(ggpubr)
library(here)
```

## A. Use the sobel approach to generate parameter values for the 4 parameters

```{r}
# Source Catm function
source(here("Catm.R"))

# generate two examples of random number from parameter distributions
np <- 1000
k_o <- rnorm(mean = 0.1, sd = 0.1 * 0.01, n = np)
k_d <- rnorm(mean = 0.7, sd = 0.7 * 0.01, n = np)
v <- rnorm(mean = 3, sd = 0.5, n = np)
height <- runif(min = 3.5, max = 5.5, n = np)

X1 <- cbind.data.frame(k_o, k_d, v, height = height)

# repeat sampling
k_o <- rnorm(mean = 0.1, sd = 0.01 * 0.1, n = np)
k_d <- rnorm(mean = 0.7, sd = 0.07 * 0.1, n = np)
v <- rnorm(mean = 3, sd = .5, n = np)
height <- runif(min = 3.5, max = 5.5, n = np)

X2 <- cbind.data.frame(k_o, k_d, v, height = height)


# Sobol object approach implemented by jansen
sens_Catm_Sobol <- sobolSalt(model = NULL, X1, X2, nboot = 100)

```

## B. Run the atmospheric conductance model for these parameters

```{r}
# run model for all parameter sets
parms <- as.data.frame(sens_Catm_Sobol$X)
colnames(parms) <- colnames(X1)
res <- pmap_dbl(parms, function(k_o, k_d, v, height) {
    Catm(v = v, height = height, k_o = k_o, k_d = k_d)
  }
)


sens_Catm_Sobol <- sensitivity::tell(sens_Catm_Sobol, res, res.names = "ga")

# main effect:  first order indices
row.names(sens_Catm_Sobol$S) <- colnames(parms)
sens_Catm_Sobol$S

# total effect indices - accounts for parameter interactions
row.names(sens_Catm_Sobol$T) <- colnames(parms)
sens_Catm_Sobol$T
```

## C. Plot conductance estimates in a way that accounts for parameter uncertainty

```{r}
# Plot conductance accounting for parameter uncertainty
both <- cbind.data.frame(parms, gs = sens_Catm_Sobol$y)

# look at overall gs sensitvity to uncertainty
ggplot(both, aes(x = gs)) +
  geom_histogram() +
  geom_vline(xintercept = mean(both$gs), col = "cyan") +
  labs(title = "Conductance Estimates",
       x = "Conductance (mm/s)",
       y = "Count"
  )
```

## D. Plot conductance estimates against windspeed use the parameter that is 2nd in terms of total effect on response

```{r}
# look at response of conductance to the two interesting variables (v and height)
ggplot(both, aes(v, gs, col = height)) +
  geom_point() +
  labs(title = "Conductance vs Windspeed",
    y = "Conductance (mm/s)", 
    x = "Windspeed (m/s)",
    color = "Vegetation Height (m)")
```

## E. Estimate the Sobel Indices for your output

```{r}
# main effect:  first order indices
row.names(sens_Catm_Sobol$S) <- colnames(parms)
sens_Catm_Sobol$S

# total effect indices - accounts for parameter interactions
row.names(sens_Catm_Sobol$T) <- colnames(parms)
sens_Catm_Sobol$T

# Both the main effect and total effect can tell us something about how the parameter influences results
print(sens_Catm_Sobol)
```

## F. Comment on what this tells you about how atmospheric conductance and its sensitivity to variation in windspeed differs in this setting as compared to the setting that we examined in class where windspeed was lower and less variable and vegetation was taller

In the assignment setting, where wind speed is higher and more varied and vegetation was shorter, conductance was more dependent on these changes and paramters (of windspeed and the vegetation height). In contrast, the in-class example - where windspeed was lower and less varied - conductance was most impacted by the kd and ko parameters. It can also be seen that these parameters, of kd and ko, did not have a relatively great impact.
